### 2013-02-10 doc/xps2013/proposal.tex
changes to intro

Irregular applications generate tasks with work, dependences, and memory
accesses that are highly sensitive to input. Classic examples of irregular
applications include branch and bound optimization, SPICE circuit simulation,
contact algorithms in car crash analysis, and network flow. Important
contemporary examples include processing large graphs in the business,
national security, machine learning, data-driven science, and social network
computing domains. For these emerging applications, fast response -- given the
sheer amount of data -- requires multinode systems. The most broadly available
multinode systems are those built from x86 commodity computing nodes
interconnected via ethernet or infiniband. Our goal is to enable scalable
performance of irregular applications on these mass market systems. Our goal
is challenging for two key reasons:

@@

In this proposal we describe a body of research that extends an early
prototype of Grappa~\cite{grappa}, a runtime system for irregular
applications. It provides several facilities: a global shared memory
abstraction, user-level threading support, and communication and
synchronization primitives. Applications targeted for Grappa are not expected
to have \emph{any} locality (although what locality that exists is favorably
exploited), but are expected to provide many thousands of concurrent threads.

Our preliminary work on Grappa shows that it runs several graph-crunching applications (classic examples of irregular behavior) efficiently on a commodity cluster.  For unbalanced tree search, Grappa is over 3X faster and shows greatly improved scalability compared to the Cray XMT; conversely, for breadth first search Grappa is 2.5X slower -- fast compared to implementations without Grappa, but obviously there is room for improvement.  Most importantly, however, for significantly less real world cost, users can \emph{add} significantly more processors to a commodity cluster than an XMT machine and use Grappa to achieve scalable performance.

The work on Grappa is still in its infancy, however. This research proposal
seeks funding to extend Grappa in the following directions:

@@

>>>>

Irregular applications generate tasks with work, dependences, and memory
accesses that are highly sensitive to input. Classic examples of irregular
applications include branch and bound optimization, SPICE circuit simulation,
contact algorithms in car crash analysis, and network flow. Important
contemporary examples include processing large graphs in the business,
national security, machine learning, data-driven science, and social network
computing domains. For these emerging applications, fast response -- given the
sheer amount of data -- requires multinode systems. The most broadly available
multinode systems are those built from x86 commodity computing nodes
interconnected via ethernet or infiniband. Our current work enables scalable
performance of irregular applications on these mass market systems,
addressing two key challenges:

@@

In this proposal we first describe our early
Grappa~\cite{grappa} prototype, a runtime system for irregular
applications. It provides several facilities: a global shared memory
abstraction, user-level threading support, and communication and
synchronization primitives. Applications targeted for Grappa are not expected
to have \emph{any} locality (although what locality that exists is favorably
exploited), but are expected to provide many thousands of concurrent threads.

Next, we present some preliminary results.  Even as an early prototype, Grappa runs several graph-crunching applications (classic examples of irregular behavior) efficiently on a commodity cluster.  For unbalanced tree search, Grappa is over 3X faster and shows greatly improved scalability compared to the Cray XMT; conversely, for breadth first search Grappa is 2.5X slower -- fast compared to implementations without Grappa, but obviously there is room for improvement.  Most importantly, however, for significantly less real world cost, users can \emph{add} significantly more processors to a commodity cluster than an XMT machine and use Grappa to achieve scalable performance.

Given these very positive early results,
we next propose to extend Grappa in the following directions:

@@

Finally, we propose to explore how the lessons learned performing this research -- specifically knowledge of what ultimately limits Grappa's performance for irregular applications using x86 processors and commodity networks with modified communication protocols -- can be applied to the design of future architectures.  Doing so will include studying the following research questions:

\being{itemize}
\item \textbf{TBD}
\end{itemize}
